---
title: "Data Management Report:- Exploratory Data Analysis"
author: "Prachi Tambe"
date: "01/12/2021"
output: pdf_document
---

# Introduction:-

## The Future Learn MOOC Dataset consisted of various Comma Seperated Files. It consists of 7 iterations of various files which contained information about students who enrolled themselves into the cyber security course over the period of two years that is from 2016 to 2018.The various data sets contained information on enrolments, surveys,leaving surveys,video streaming information, etc. The data could be used to perform various analysis to help the management to come with new strategies and by giving an detailed insight to the observations so that the business strategies related to it can take a step-up and upscale their businesses accordingly.

## After careful observation in leaving surveys of the students it was noted that there were quite a few conclusions that could be drawn based on the analysis.
## The exploratory data analysis in this field could help the management recover the loss due to students failing to complete their courses. 
## To start with the EDA Project Template proved to be handy.
## The data was extracted and the EDA could be performerd in no time without any hasle.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = TRUE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
```

```{r ProjectTemplate, include=FALSE}
library(ProjectTemplate)
load.project()
```
# Starting the EDA

```{r}
##install.packages("readr")
library(readr)
```
## The EDA was to be carried on the leaving survey data. Hence it was loaded and stored as data frames. It only makes sense the carry out the analysis on the latest survey. Hence last four leaving survey data was used in this analysis. The data from fourth and fifth iteration was used for the first cycle and the data from sixth and seventh iteration was used in the second cycle of the exploratory data analysis.

```{r}
leaving_survey_responses4 <- read_csv("/Users/prachitambe/Desktop/DM/data/cyber-security-4_leaving-survey-responses.csv")
leaving_survey_responses5 <- read_csv("/Users/prachitambe/Desktop/DM/data/cyber-security-5_leaving-survey-responses.csv")
leaving_survey_responses6 <- read_csv("/Users/prachitambe/Desktop/DM/data/cyber-security-6_leaving-survey-responses.csv")
leaving_survey_responses7 <- read_csv("/Users/prachitambe/Desktop/DM/data/cyber-security-7_leaving-survey-responses.csv")
```
## The fourth and fifth iteration of leaving survey data was megered to perform first cycle in eda.
```{r}
leaving_survey_responses_data1 <- rbind(leaving_survey_responses4,leaving_survey_responses5)
dim(leaving_survey_responses_data1)
```
## The sixth and seventh iteration of leaving survey data was megered to perform second cycle in eda.
```{r}
leaving_survey_responses_data2 <- rbind(leaving_survey_responses6,leaving_survey_responses7)
dim(leaving_survey_responses_data2)
```
## After cleaning and merging the data sets the eda can now begin. 
## To begin with, the reasons given by the students for leaving the course might provide us with the insights of the real situation.To know the main reason a the data was visualised in the form of bar plots using the library ggplot.
```{r}
library(ggplot2)
ggplot(data = leaving_survey_responses_data1, mapping = aes(x =leaving_reason, fill = leaving_reason))+geom_bar() + labs (title = "Comparision between the reasons that made the students leave the course", x = "Leaving Reasons")+theme(axis.text.x = element_blank())
```

## Hence we can be concluded that the maximum students who choosed to leave the course had some other reason than not having enough time irrelevance, course being easy or hard or unexpected, time consuming. So there is nothing that could be done much here. The second main reason was not having enough time. The same thing was done in the second cycle to be sure of the conclusions.

```{r}
library(ggplot2)
   ggplot(data = leaving_survey_responses_data2, mapping = aes(x =leaving_reason, fill = leaving_reason))+geom_bar() + labs (title = "Comparision betwwen the reasons the students are leaving the course", x = "Leaving Reasons")+theme(axis.text.x = element_blank())
```

## In the second cycle it was clear that the maximum student were struggling with time and choose to quit.
## To overcome this problem from business perspective and keeping the credibility of the course unharmed it would be advisible to make changes in the course work. Make it more understandable in less time. The course video lectures should be trimmed and unneccessary content should be removed tommake it more engaging.
 
## The data was cleaned for further analysis.
```{r}
#Cleaning the data
clean_leaving_survey_responses1 <- na.omit(leaving_survey_responses_data1)

dim(clean_leaving_survey_responses1)
```

```{r}
#Cleaning the data
clean_leaving_survey_responses2 <- na.omit(leaving_survey_responses_data2)

dim(clean_leaving_survey_responses2)
```

## To understand more aabout the background of the students who choose to quit the course the enrolment data was handy and proved to be a lot helpful to draw useful conclusions.
## Similar to the leaving  data the fourth, fifth ,sixth and seventh iteration of the data was used for ease of the process.
## Hence the the was loaded and stored.
```{r}
enrolments4 <- read_csv("/Users/prachitambe/Desktop/DM/data/cyber-security-4_enrolments.csv")
enrolments5 <- read_csv("/Users/prachitambe/Desktop/DM/data/cyber-security-5_enrolments.csv")
enrolments6 <- read_csv("/Users/prachitambe/Desktop/DM/data/cyber-security-6_enrolments.csv")
enrolments7 <- read_csv("/Users/prachitambe/Desktop/DM/data/cyber-security-7_enrolments.csv")
```
## The fourth and fifth iterations were combined to perform first cycle.
```{r}
enrolments1 <- rbind(enrolments4,enrolments5)
dim(enrolments1)
```
## The sixth and seventh iterations were combined to perform second cycle.
```{r}
enrolments2 <- rbind(enrolments6,enrolments7)
dim(enrolments2)
```
## Irrelevant columns were removed to make the data more relevant and easy to access.
```{r}
##Deleting Unnecessary columns
colnames(enrolments1)
enrolments_new1 <- enrolments1[-c(2, 3:6, 13)]
dim(enrolments_new1)
```

```{r}
##Deleting Unnecessary columns
colnames(enrolments2)
enrolments_new2 <- enrolments2[-c(2, 3:6, 13)]
dim(enrolments_new2)
```
## To blank rows in the data were omitted to improve the efficiency of eda.
```{r}
##Cleaning the data
clean_enrolments1 <- na.omit(enrolments_new1)
dim(clean_enrolments1)
```

```{r}
##Cleaning the data
clean_enrolments2 <- na.omit(enrolments_new2)
dim(clean_enrolments1)
```
## Finally the enrolment data was merged with leaving responses dta using the common column learner_id.
```{r}
newdata1 <- merge(clean_leaving_survey_responses1 ,clean_enrolments1, by.x = "learner_id")

```

```{r}
newdata2 <- merge(clean_leaving_survey_responses2 ,clean_enrolments2, by.x = "learner_id")

```
## The new data was processed and stored. The information on all the students who choosed to quit the course was obtained so that we can draw more relevant conclusions on the student choices.
## The unknown information was removed to ease the eda procress for first cycle.
```{r}
clean_newdata1 <- newdata1[newdata1$employment_status!="Unknown",]
clean_newdata1 <- newdata1[newdata1$age_range!="Unknown",]
clean_newdata1 <- newdata1[newdata1$highest_education_level!="Unknown",]
```
## The unknown information was removed to ease the eda procress for second cycle.
```{r}
clean_newdata2 <- newdata2[newdata2$employment_status!="Unknown",]
clean_newdata2 <- newdata2[newdata2$age_range!="Unknown",]
clean_newdata2 <- newdata2[newdata2$highest_education_level!="Unknown",]
```
## To begin with ,the Employment Status of the students who choosed to quit was visualised in the form of bar plots using ggplot.
## Begining with the first cycle
```{r}
ggplot(data = clean_newdata1, mapping = aes(x =employment_status, fill = employment_status))+geom_bar() + labs (title = "Employment status of the students are leaving the course", x = "Employment Status")+theme(axis.text.x = element_blank())
```

## Here it can be clearly concluded that the students who were retired and working full time were not able to complete the course. It is quite obvious that the employed students might not be able to complete it because of lack of time.
## Second Cycle was processed to check if there is a similar result
```{r}
ggplot(data = clean_newdata2, mapping = aes(x =employment_status, fill = employment_status))+geom_bar() + labs (title = "Employment status of the students are leaving the course", x = "Employment Status")+theme(axis.text.x = element_blank())
```

## The maximum students who choosed to quit were both working and non working students.
## Since the working students might not find much time in hands they must have choosed to quick since it was the most common reason 
## The non working students could be any one between the range less than 18 or more that 65 which might have made the course too hard or easy or irrelevant.
## The business team should clarify the time and engagement requirments in the course so that students have an idea as to how time consuming it could get so thta they can make an assumption whether they could do it or not. The second option is to reduce the course period to make it kess time consuming or generating a whole new content for working people.
## The analysis was continued on the basis of age the range to see if the non working category could fall under the age range which is slightly higher than normal

## First Cycle:-
```{r}
ggplot(data = clean_newdata1, mapping = aes(x =age_range, fill = age_range))+geom_bar() + labs (title = "Age range of the students are leaving the course", x = "Age range")+theme(axis.text.x = element_blank())
```

## Clearly the age range of students from 46 to more than 65 were more likely  to quit the course as compared to other.
## Analysis was reconfirmed by second cycle.
```{r}
 ggplot(data = clean_newdata2, mapping = aes(x =age_range, fill = age_range))+geom_bar() + labs (title = "Age range of the students are leaving the course", x = "Age range")+theme(axis.text.x = element_blank()) 
```
## Here it can be clearly  noted that the students from age range of 46 to 65 were more likely to quit the course.
## Hence we can conclude that the students from age more than 46 are more likely to quit.
## The Business team should consider this information and provide proper information to the students who wish to enroll and are above 46 so that they can make their decision accordingly hence their would be a less chance they quit overall maintaining the the authenticity and crdibility of the course.

##The analysis was continued by making conclusions on decisions made by students on the basis of their higher education.
##First Cycle
```{r}
ggplot(data = clean_newdata1, mapping = aes(x = highest_education_level, fill = highest_education_level))+geom_bar() + labs (title = "Highest education of the students are leaving the course", x = "Highest Education Level")+theme(axis.text.x = element_blank())
```

##Here it can be clearly seen that the students who were under graduate are most likely to quit followed by postgraduate students.
##Performing second cycle to reconfirm.
```{r}
ggplot(data = clean_newdata2, mapping = aes(x = highest_education_level, fill = highest_education_level))+geom_bar() + labs (title = "Highest education of the students are leaving the course", x = "Highest education Level")+theme(axis.text.x = element_blank())
```

##Here it can be clearly seen that the students who were under graduate are most likely to quit followed by postgraduate students.
##The main reason could be being employed.
##The main reason that the university students can be concluded if we know the emplyment status of the undergrad and post gard students.

##Performing the eda to find out the employment status of the Students based on their highest education level

##First Cycle
```{r}
ggplot(data = clean_newdata1, mapping = aes(x = highest_education_level, fill = employment_status))+geom_bar() + labs (title = "Highest education of the students are leaving the course", x = "Highest Education Level")+theme(axis.text.x = element_blank())
```
## Here we can clearly see that the undergrad students who are more likely to quit were mostly working or looking for work or retired, hence we can say that they fall short of time and end up quiting. The post grad students are mostly retired or working full time.

##Second Cycle
```{r}
ggplot(data = clean_newdata2, mapping = aes(x = highest_education_level, fill = employment_status))+geom_bar() + labs (title = "Highest education of the students are leaving the course", x = "Highest Education Level")+theme(axis.text.x = element_blank())
```

##Here we can see that the post grad students who are likely to quit are working full time or are retired and the under grads who are likely to quit the course are employed or atleast looking for work.

# Conclusion:-
## From the above made conclusions it can be concluded that the students who are likely ti quit fall under the age range of above 46 - 65 and are atleast graduated or post graduates and fall under working category. The course should be marketed keeping all this things in considerations. The working category students should bw aware of the fact that the course may consume slightly more time than expected. To avoid the backing off the course must be curated keeping this students in mind. They should be lessmtime consuming, on point and not repetitive.
## This will probably help the institute to improve its credibility and prove to benificial to all.

